Manuscript Draft

### 2019-01-08 manuscript.Rmd
update rmds

```{r grid-wilcox-res, echo = FALSE}
ccle_signif_g2p_grid <- readRDS("./data_munging/rds/ccle_signif_g2p_grid.rds")
ctrp_signif_g2p_grid <- readRDS("./data_munging/rds/ctrp_signif_g2p_grid.rds")
gdsc_signif_g2p_grid <- readRDS("./data_munging/rds/gdsc_signif_g2p_grid.rds")
```

@@

@@

@@

In the `r nrow(g2p_filt)` level-A G2P interpretations, there was a wide variation in the specificity of cancer description, ranging from “Waldenström macroglobulinemia” and “hyper eosinophilic advanced syndrome” on the more detailed end to “cancer” on the broad end. Similarly, the drug screen datasets had a wide range in lineage specificity, e.g. CCLE, GTRP, and GDSC all had lineages labeled “leukemia” and “T-cell childhood acute lymphocytic leukemia” and the CRISPR dataset had cell lines labeled “Epstein-Barr virus-related Burkitt lymphoma” and “lymphoma.” To more easily conduct comparative analyses in a cancer-specific manner, we developed a more general lineage grouping method derived from the Human Disease Ontology identity codes (DOIDs, http://disease-ontology.org/) [@schriml2019] assigned to each cell line, yielding `r length(unique(ccl_converter$group_general_lineage_name))` unique lineage groupings. This custom lineage grouping was included in our harmonized cancer cell line database (see METHODS).

@@

At an $\alpha =  0.05$, no DGLs were identified in the CCLE screen when using mutation status as a correlate (n = 7), but CTRP and GDSC identified six and five, respectively, in skin cancer, lung cancer, and cancer. These significant associations did not hold after Bonferroni correction in CTRP ($n = 48$, $\alpha_{adjusted} = 0.001$) or GDSC ($n = 39$, $\alpha_{adjusted} = 0.001$).

When using gene expression as the genomic feature, three CCLE DGLs had significant correlations at an adjusted $\alpha =  0.006$ ($n = 9$). Two of these three DGLs were also significantly correlated with copy number ($n = 9$, $\alpha_{adjusted} = 0.006$).

In CTRP, AUC of 19 DGLs significantly correlated with gene expression, 12 of which were significant after correction ($n = 61$, $\alpha_{adjusted} = 0.0008$). Copy number correlations were significant for 12 DGLs, with three DGLs significant post-correction ($n = 61$, $\alpha_{adjusted} = 0.0008$).

AUC of 16 DGLs in GDSC significantly correlated with gene expression. At an adjusted $\alpha$ of 0.0009 ($n = 53$) eight of them were significant. Fewer DGLs were significant when correlating AUC with copy number with 6 significant at an $\alpha$ of 0.05 and 1 significant at an adjusted $\alpha$ of 0.0009 ($n = 53$).

Interestingly, when we test the effect of mutation status on drug response for the DGLs derived from all pairwise combinations of G2P drugs and genes… (*Figure ##1*).

>>>>

```{r grid-dg-wilcox, echo = FALSE}
ccle_signif_g2p_grid <- readRDS("./data_munging/rds/ccle_signif_g2p_grid.rds")
ctrp_signif_g2p_grid <- readRDS("./data_munging/rds/ctrp_signif_g2p_grid.rds")
gdsc_signif_g2p_grid <- readRDS("./data_munging/rds/gdsc_signif_g2p_grid.rds")
```

@@

**Cut this analysis since we are comparing distributions of p-values?**

@@

```{r grid-dgl-wilcox, echo = FALSE}
ccle_signif_g2p_lineage <- readRDS("./data_munging/rds/ccle_signif_g2p_lineage.rds")
ccle_signif_grid_lineage <- readRDS("./data_munging/rds/ccle_signif_grid_lineage.rds")

ctrp_signif_g2p_lineage <- readRDS("./data_munging/rds/ctrp_signif_g2p_lineage.rds")
ctrp_signif_grid_lineage <- readRDS("./data_munging/rds/ctrp_signif_grid_lineage.rds")

gdsc_signif_g2p_lineage <- readRDS("./data_munging/rds/gdsc_signif_g2p_lineage.rds")
gdsc_signif_grid_lineage <- readRDS("./data_munging/rds/gdsc_signif_grid_lineage.rds")
```

@@

In the `r nrow(g2p_filt)` level-A G2P interpretations, there was a wide variation in the specificity of cancer description, ranging from “Waldenström macroglobulinemia” and “hyper eosinophilic advanced syndrome” on the more detailed end to “cancer” on the broad end. Similarly, the drug screen datasets had a wide range in lineage specificity, e.g. CCLE, GTRP, and GDSC all had lineages labeled “leukemia” and “T-cell childhood acute lymphocytic leukemia” and the CRISPR dataset had cell lines labeled “Epstein-Barr virus-related Burkitt lymphoma” and “lymphoma.” To more easily conduct comparative analyses in a cancer-specific manner, we developed a more general lineage grouping method derived from the Human Disease Ontology identity codes (DOIDs, http://disease-ontology.org/) [@schriml2019] assigned to each cell line, yielding `r length(unique(ccl_converter$group_general_lineage_name))` unique lineage groupings. This custom lineage grouping was included in our harmonized cancer cell line database (see Methods).

@@

Prior to correction for an FDR of 5%, no DGLs were identified in the CCLE screen when using mutation status as a correlate ($n =$ `r nrow(ccle_signif_g2p_lineage)`, `r length(unique(ccle_data_g2p_dgl$Drug_Gene_Lin)) - nrow(ccle_signif_g2p_lineage)` DGLs were not analyzed due to lack of mutation annotations in *UGT1A1* in "cancer" lineages), but CTRP ($n =$ `r nrow(ctrp_signif_g2p_lineage)`) and GDSC ($n =$ `r nrow(gdsc_signif_g2p_lineage)`) identified `r nrow(filter(ctrp_signif_g2p_lineage, !is.na(p.signif)))` and `r nrow(filter(gdsc_signif_g2p_lineage, !is.na(p.signif)))`, respectively, in skin cancer, lung cancer, and cancer. After Benjamini-Hochberg FDR correction, none of the datasets yielded significant DGL interactions.

**BAD ANALYSIS, keep for reference:** These significant associations did not hold after Bonferroni correction in CTRP ($n = 48$, $\alpha_{adjusted} = 0.001$) or GDSC ($n = 39$, $\alpha_{adjusted} = 0.001$). When using gene expression as the genomic feature, three CCLE DGLs had significant correlations at an adjusted $\alpha =  0.006$ ($n = 9$). Two of these three DGLs were also significantly correlated with copy number ($n = 9$, $\alpha_{adjusted} = 0.006$). In CTRP, AUC of 19 DGLs significantly correlated with gene expression, 12 of which were significant after correction ($n = 61$, $\alpha_{adjusted} = 0.0008$). Copy number correlations were significant for 12 DGLs, with three DGLs significant post-correction ($n = 61$, $\alpha_{adjusted} = 0.0008$). AUC of 16 DGLs in GDSC significantly correlated with gene expression. At an adjusted $\alpha$ of 0.0009 ($n = 53$) eight of them were significant. Fewer DGLs were significant when correlating AUC with copy number with 6 significant at an $\alpha$ of 0.05 and 1 significant at an adjusted $\alpha$ of 0.0009 ($n = 53$).

### 2019-01-08 manuscript.tex
update rmds

@@

In the 894 level-A G2P interpretations, there was a wide variation in
the specificity of cancer description, ranging from \enquote{Waldenström
macroglobulinemia} and \enquote{hyper eosinophilic advanced syndrome} on
the more detailed end to \enquote{cancer} on the broad end. Similarly,
the drug screen datasets had a wide range in lineage specificity,
e.g.~CCLE, GTRP, and GDSC all had lineages labeled \enquote{leukemia}
and \enquote{T-cell childhood acute lymphocytic leukemia} and the CRISPR
dataset had cell lines labeled \enquote{Epstein-Barr virus-related
Burkitt lymphoma} and \enquote{lymphoma.} To more easily conduct
comparative analyses in a cancer-specific manner, we developed a more
general lineage grouping method derived from the Human Disease Ontology
identity codes (DOIDs, \url{http://disease-ontology.org/}) (Schriml et
al., 2019) assigned to each cell line, yielding 27 unique lineage
groupings. This custom lineage grouping was included in our harmonized
cancer cell line database (see METHODS).

@@

At an \(\alpha = 0.05\), no DGLs were identified in the CCLE screen when
using mutation status as a correlate (n = 7), but CTRP and GDSC
identified six and five, respectively, in skin cancer, lung cancer, and
cancer. These significant associations did not hold after Bonferroni
correction in CTRP (\(n = 48\), \(\alpha_{adjusted} = 0.001\)) or GDSC
(\(n = 39\), \(\alpha_{adjusted} = 0.001\)).

When using gene expression as the genomic feature, three CCLE DGLs had
significant correlations at an adjusted \(\alpha = 0.006\) (\(n = 9\)).
Two of these three DGLs were also significantly correlated with copy
number (\(n = 9\), \(\alpha_{adjusted} = 0.006\)).

In CTRP, AUC of 19 DGLs significantly correlated with gene expression,
12 of which were significant after correction (\(n = 61\),
\(\alpha_{adjusted} = 0.0008\)). Copy number correlations were
significant for 12 DGLs, with three DGLs significant post-correction
(\(n = 61\), \(\alpha_{adjusted} = 0.0008\)).

AUC of 16 DGLs in GDSC significantly correlated with gene expression. At
an adjusted \(\alpha\) of 0.0009 (\(n = 53\)) eight of them were
significant. Fewer DGLs were significant when correlating AUC with copy
number with 6 significant at an \(\alpha\) of 0.05 and 1 significant at
an adjusted \(\alpha\) of 0.0009 (\(n = 53\)).

Interestingly, when we test the effect of mutation status on drug
response for the DGLs derived from all pairwise combinations of G2P
drugs and genes\ldots{} (\emph{Figure \#\#1}).

>>>>

\textbf{Cut this analysis since we are comparing distributions of
p-values?}

@@

In the 894 level-A G2P interpretations, there was a wide variation in
the specificity of cancer description, ranging from \enquote{Waldenström
macroglobulinemia} and \enquote{hyper eosinophilic advanced syndrome} on
the more detailed end to \enquote{cancer} on the broad end. Similarly,
the drug screen datasets had a wide range in lineage specificity,
e.g.~CCLE, GTRP, and GDSC all had lineages labeled \enquote{leukemia}
and \enquote{T-cell childhood acute lymphocytic leukemia} and the CRISPR
dataset had cell lines labeled \enquote{Epstein-Barr virus-related
Burkitt lymphoma} and \enquote{lymphoma.} To more easily conduct
comparative analyses in a cancer-specific manner, we developed a more
general lineage grouping method derived from the Human Disease Ontology
identity codes (DOIDs, \url{http://disease-ontology.org/}) (Schriml et
al., 2019) assigned to each cell line, yielding 27 unique lineage
groupings. This custom lineage grouping was included in our harmonized
cancer cell line database (see Methods).

@@

Prior to correction for an FDR of 5\%, no DGLs were identified in the
CCLE screen when using mutation status as a correlate (\(n =\) 7, 2 DGLs
were not analyzed due to lack of mutation annotations in \emph{UGT1A1}
in \enquote{cancer} lineages), but CTRP (\(n =\) 48) and GDSC (\(n =\)
39) identified 6 and 5, respectively, in skin cancer, lung cancer, and
cancer.

\textbf{BAD ANALYSIS, keep for reference:} These significant
associations did not hold after Bonferroni correction in CTRP
(\(n = 48\), \(\alpha_{adjusted} = 0.001\)) or GDSC (\(n = 39\),
\(\alpha_{adjusted} = 0.001\)). When using gene expression as the
genomic feature, three CCLE DGLs had significant correlations at an
adjusted \(\alpha = 0.006\) (\(n = 9\)). Two of these three DGLs were
also significantly correlated with copy number (\(n = 9\),
\(\alpha_{adjusted} = 0.006\)). In CTRP, AUC of 19 DGLs significantly
correlated with gene expression, 12 of which were significant after
correction (\(n = 61\), \(\alpha_{adjusted} = 0.0008\)). Copy number
correlations were significant for 12 DGLs, with three DGLs significant
post-correction (\(n = 61\), \(\alpha_{adjusted} = 0.0008\)). AUC of 16
DGLs in GDSC significantly correlated with gene expression. At an
adjusted \(\alpha\) of 0.0009 (\(n = 53\)) eight of them were
significant. Fewer DGLs were significant when correlating AUC with copy
number with 6 significant at an \(\alpha\) of 0.05 and 1 significant at
an adjusted \(\alpha\) of 0.0009 (\(n = 53\)).
